---
title: "COVID-19 Data Analysis"
author: "Hamish Spalding"
date: "16/03/2020"
output: 
  word_document:
    reference_docx: HWordStyleReference.docx
params:
  date: !r Sys.Date()
  author: Hamish

---

```{r libs, eval=TRUE, echo=FALSE}
library(pacman)
p_load(dplyr,readr, glue, lubridate, gridExtra, kableExtra, tidyr, ggplot2, cowplot, forcats,stringr)
p_load(gganimate)
```

```{r global_options, include=FALSE}
knitr::opts_chunk$set(fig.width=6, fig.height=4, fig.path='Figs/',
                      echo=FALSE, include = FALSE, warning=FALSE, message=FALSE)
```

## Data Source

This report was produced on `r params$date` by `r params$author`. It relies on data was pulled from the the 2019 Novel Coronavirus Visual Dashboard operated by the Johns Hopkins University Center for Systems Science and Engineering (JHU CSSE). A complete list of sources is available at https://github.com/CSSEGISandData/COVID-19 .

The licence for the data is that JHU CSSE retains copyright and does not permit the use of the data for commercial purposes.

Dates on charts are as at date of running the report. Data is 1 day behind this due to US time zone.

```{r}
# Raw paths to .csv on github:
#https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_time_series/time_series_19-covid-Confirmed.csv
#https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_time_series/time_series_19-covid-Deaths.csv
#https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_time_series/time_series_19-covid-Recovered.csv

url = "https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_time_series/time_series_19-covid-Confirmed.csv"
destfile = "time_series_19-covid-Confirmed.csv"
confirmed <- download.file(url, destfile)

url = "https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_time_series/time_series_19-covid-Deaths.csv"
destfile = "time_series_19-covid-Deaths.csv"
download.file(url, destfile)

url = "https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_time_series/time_series_19-covid-Recovered.csv"
destfile = "time_series_19-covid-Recovered.csv"
download.file(url, destfile)

```



```{r, warning=FALSE}
checks <- read_csv("time_series_19-covid-Confirmed.csv") %>%
  rename(country=`Country/Region`)  %>%
  rename(state=`Province/State`)  %>%
  mutate(country_state = paste(country, state, sep = "_")) %>%
  select(-c(state, Lat, Long))%>%  
  filter(country=="Australia") 
#dim(checks)
#names(checks) # to get date
#but last column is checks[,ncol(checks)-1]
#Sys.Date() # "2020-03-25"


checks.australia.state <- select(checks,country_state,names(checks[,ncol(checks)-1]))
checks.australia.total <- select(checks,country_state,names(checks[,ncol(checks)-1])) %>% select(names(checks[,ncol(checks)-1])) %>% sum()
checks.australia.state
checks.australia.total

```


```{r}
## Read In and Combine Data

get.data <- function(csvfile="time_series_19-covid-Confirmed.csv",
         group_label="Confirmed"){                            
  df <- read_csv(csvfile) %>%
  rename(country=`Country/Region`)  %>%
  rename(state=`Province/State`)  %>%
  mutate(country_state = paste(country, state, sep = "_")) %>%
  select(-c(state, Lat, Long, country_state))%>%  #%>%filter(country=="Australia")
  pivot_longer(-country, names_to = "date", values_to = "n") %>%
  mutate(date = as.Date(date, format = "%m/%d/%y")) %>% 
  mutate(group=group_label) %>%
  group_by(country,date, group) %>% summarize(n = sum(n))
  df
}

get.data.state <- function(csvfile="time_series_19-covid-Confirmed.csv",
         group_label="Confirmed"){
  df <- read_csv(csvfile) %>%
    rename(country=`Country/Region`)  %>%
    rename(state=`Province/State`)  %>%
    mutate(country_state = paste(country, state, sep = "_")) %>%
    select(-c(state, Lat, Long, country))%>%
  #select(-c(state, Lat, Long, country_state))%>%  #%>%filter(country=="Australia")
    pivot_longer(-country_state, names_to = "date", values_to = "n") %>%
    mutate(date = as.Date(date, format = "%m/%d/%y")) %>%
    mutate(group=group_label) %>%
    group_by(country_state,date, group) %>% summarize(n = sum(n))
  df
}

confirmed <- get.data("time_series_19-covid-Confirmed.csv",group_label="Confirmed")
deaths <- get.data("time_series_19-covid-Deaths.csv",group_label="Deaths")
recovered <- get.data("time_series_19-covid-Recovered.csv",group_label="Recovered")

confirmed.state <- get.data.state("time_series_19-covid-Confirmed.csv",group_label="Confirmed")
deaths.state <- get.data.state("time_series_19-covid-Deaths.csv",group_label="Deaths")
recovered.state <- get.data.state("time_series_19-covid-Recovered.csv",group_label="Recovered")


#confirmed <- read_csv("time_series_19-covid-Confirmed.csv") %>%
#  rename(country=`Country/Region`)  %>%
#  rename(state=`Province/State`)  %>%
#  mutate(country_state = paste(country, state, sep = "_")) %>%
#  #select(-c(state, Lat, Long))%>%  
#  select(-c(state, Lat, Long, country_state))%>%  #%>%filter(country=="Australia")
#  pivot_longer(-country, names_to = "date", values_to = "n") %>%
#  mutate(date = as.Date(date, format = "%m/%d/%y")) %>% 
#  mutate(group="Confirmed")#

all <- rbind(confirmed, deaths, recovered)
all.state <- rbind(confirmed.state,deaths.state, recovered.state)

aus <- filter(all, country=="Australia")

all.state <- rbind(confirmed.state,deaths.state,recovered.state) 

aus.state <- filter(all.state, str_detect(country_state, "Australia"))

```

## Plots


## Australia vs other Countries that Contained Well

```{r include=TRUE, warning=FALSE}
all %>% 
  filter(date >= "2020-03-01" & date <= Sys.Date()) %>%
  #filter(group == "Confirmed") %>% # choose: "Confirmed" "Deaths"    "Recovered"
  filter(country %in% c("Australia","Singapore")) %>%
           #, "Iran", "Italy", "US", "Singapore")) %>% # choose country
  ggplot(aes(x = date, y = n)) +
  geom_line(aes(color = country, linetype = country)) +
  theme_light() +
  scale_y_log10() +
  facet_grid(group~.)
```

## Several Countries in Same chart

```{r include=TRUE, warning=FALSE}
all %>% 
  filter(date >= "2020-02-20" & date <= Sys.Date()) %>%
  #filter(date >= as.Date("2020-03-07") & date <= as.Date("2020-03-19")) %>%
  filter(group == "Confirmed") %>% # choose: "Confirmed" "Deaths"    "Recovered"
  filter(country %in% c("Australia", "Singapore", "United Kingdom", "US")) %>% # choose country
  ggplot(aes(x = date, y = n)) +
  geom_line(aes(color = country, linetype = country)) + 
  scale_color_manual(values = c("black", "steelblue", "red", "blue")) +
  theme_minimal_vgrid() +
  scale_y_log10()

```

## Several states in Same chart

```{r include=TRUE, warning=FALSE}
aus.state %>% 
  filter(date >= "2020-02-20" & date <= Sys.Date()) %>%
  filter(group == "Confirmed") %>% # choose: "Confirmed" "Deaths"    "Recovered"
  ggplot(aes(x = date, y = n)) +
  geom_line(aes(color = country_state, linetype = country_state)) +
  theme_minimal_vgrid() +
  scale_y_log10()
  #scale_color_manual(values = c("black", "steelblue", "red", "blue"))
  
```


```{r include=TRUE, warning=FALSE}
dateforplot <- params$date
dateforplot <- as.character(dateforplot)
aus.state %>% 
  filter(date >= "2020-03-01" & date <= Sys.Date()) %>%
  filter(group == "Confirmed") %>% # choose: "Confirmed" "Deaths"    "Recovered"
  ggplot(aes(x = date, y = n)) +
  geom_line(aes(color = country_state, linetype = country_state)) + 
  facet_grid(group~.) +
  labs(title = "Total Daily Confirmed Cases",
           subtitle = paste0(glue(dateforplot)),
           x = "Date", y = "Cases") +
  theme_minimal_vgrid()

aus.state.fit <- aus.state %>% 
  filter(date >= "2020-03-01" & date <= Sys.Date()) %>%
  filter(country_state=="Australia_New South Wales") %>% #
  filter(group == "Confirmed") # choose: "Confirmed" "Deaths"    "Recovered"

exponential.model <- lm(log(aus.state.fit$n)~ aus.state$date)
summary(exponential.model)

y <- aus.state.fit$country_state
x <- aus.state.fit$date
Estimate = lm(y ~ x)
logEstimate = lm(y ~ log(x))

plot(x,predict(Estimate),type='l',col='blue')
lines(x,predict(logEstimate),col='red')
points(x,y)



  ggplot(aes(x = date, y = n)) +
  geom_line(aes(color = country_state, linetype = country_state))

```


## Bin

```{r plots, eval=FALSE, include=FALSE}

dateforplot <- params$date
dateforplot <- as.character(dateforplot)
#glue(dateforplot)
#paste0(glue(dateforplot))

#class(data$date) # chr - need to convert this to a date class
#class(data$Confirmed) # numeric
## counts for the whole world 
plot_confirmed <- confirmed %>%# select(-country) %>%
  #filter(country=="Australia") %>%
  #mutate(date = as.Date(date, format = "%m/%d/%y")) %>% # converts to a date class
  group_by(date,country) %>%
  summarise(Confirmed = sum(n)) %>%
  ggplot(aes(x = date, y = Confirmed)) + geom_line() + geom_point() +
      geom_col(fill = "orange") +
      labs(title = "Total Daily Confirmed Cases",
           #subtitle = "19 March 2020", 
           subtitle = paste0(glue(dateforplot)),
           x = "Date", y = "Cases") +
  theme_minimal_vgrid()
  
plot_deaths <- deaths %>%# select(-country) %>%
  select(-group)%>%
  #filter(country=="Australia") %>%
  group_by(date) %>%
  summarise(Deaths = sum(Deaths)) %>%
  mutate(date = as.Date(date, format = "%m/%d/%y")) %>% # converts to a date class
  ggplot(aes(x = date, y = Deaths)) + geom_line() +
      #geom_col(fill = "orange") +
      labs(title = "Total Daily Deaths",
           subtitle = paste0(glue(dateforplot)),
           x = "Date", y = "Deaths") +
  theme_light() + theme_linedraw()

plot_recovered <- recovered %>%# select(-country) %>%
  #filter(country=="Australia") %>%
  select(-group)%>%
  group_by(date) %>%
  summarise(Recovered = sum(Recovered)) %>%
  mutate(date = as.Date(date, format = "%m/%d/%y")) %>% # converts to a date class
  ggplot(aes(x = date, y = Recovered)) + geom_line() +
      geom_col(fill = "orange") +
      labs(title = "Total Daily Recoveries",
           subtitle = paste0(glue(dateforplot)),
           x = "Date", y = "Recoveries") +
  theme_light()


plot_grid(plot_confirmed, plot_deaths, plot_recovered, nrow = 3,
            # labels = c('Confirmed', 'Deaths', 'Recovered'), 
          label_size = 12)

plot_confirmed
plot_deaths
plot_recovered


rbind(recovered, confirmed, deaths)
plot_all <- rbind(recovered, confirmed, deaths) %>%# select(-country) %>%
  #filter(country=="Australia") %>%
  
  group_by(date) %>%
  summarise(Recovered = sum(Recovered)) %>%
  mutate(date = as.Date(date, format = "%m/%d/%y")) %>% # converts to a date class
  ggplot(aes(x = date, y = Recovered)) + geom_line() +
      geom_col(fill = "orange") +
      labs(title = "Total Daily Recoveries",
           subtitle = paste0(glue(dateforplot)),
           x = "Date", y = "Recoveries") +
  theme_light()


```

```{r eval=FALSE, include=FALSE}
pie_chart <- function(df, group_var, title) {
  group_var <- enquo(group_var) # to quote
  title <- enquo(title)         # to quote
  df %>%
    select(!!group_var) %>% 
    group_by(!!group_var) %>% 
    summarise(Count = n()) %>%
    mutate(per=Count/sum(Count)) %>%
    ggplot(aes(x="", y=Count, fill=!!group_var)) + # !!to unquote
    geom_bar(stat="identity", width=1, color="white") +
    geom_text(aes(label = paste(round(Count / sum(Count) * 100, 1), "%")),
              position = position_stack(vjust = 0.5)) +
    coord_polar("y", start=0) + theme_void() +
    #labs(title = title,subtitle = paste0(glue(dateforplot))) +
    scale_fill_brewer(palette="Oranges")
}

pie_chart(x,group_var = country, title = "Top 10 Confirmed by Country")

# pie of Confirmed by country
pie_confirmed <-  confirmed %>% # "country"   "date"      "Confirmed"
  filter(date == max(date)) %>%
  group_by(country) %>%
  summarise(Confirmed_Total=sum(Confirmed)) %>% 
  mutate(ranking = dense_rank(desc(Confirmed_Total))) %>%
  filter(ranking <= 8) %>%
  mutate(per=Confirmed_Total/sum(Confirmed_Total)) %>%

  ggplot(aes(x="", y=Confirmed_Total, fill=fct_reorder(country, per, .desc = TRUE))) + # !!to unquote
  geom_bar(stat="identity", width=1, color="white") +
  geom_text(aes(label = paste(round(Confirmed_Total / sum(Confirmed_Total) * 100, 1), "%")),
           position = position_stack(vjust = 0.5)) +
  coord_polar("y", start=0) + theme_void() +
  #labs(title = title,subtitle = paste0(glue(dateforplot))) +
  scale_fill_brewer(palette="Oranges")
  

pie_chart2 <- function(df, group_var, title) {
  group_var <- enquo(group_var) # to quote confirmed
  title <- enquo(title)         # to quote
  df %>%
    filter(date == max(date)) %>%
    group_by(!!group_var) %>%
    summarise(group_var=sum(!!group_var)) %>% 
    mutate(ranking = dense_rank(desc(!!group_var))) %>%
    filter(ranking <= 8) %>%
    mutate(per=!!group_var/sum(!!group_var)) %>%
    
    ggplot(aes(x="", y=!!group_var, fill=fct_reorder(!!group_var, per, .desc = TRUE))) + # !!to unquote
    geom_bar(stat="identity", width=1, color="white") +
    geom_text(aes(label = paste(round(Confirmed_Total / sum(Confirmed_Total) * 100, 1), "%")), 
              position = position_stack(vjust = 0.5)) +
    
  coord_polar("y", start=0) + theme_void() +
  #labs(title = title,subtitle = paste0(glue(dateforplot))) +
  scale_fill_brewer(palette="Oranges")
}


pie_chart2(confirmed,group_var = country, title = "Top 10 Confirmed by Country")

  
  
  
  
  







# pie of Deaths by country
pie_deaths <-  deaths %>% filter(date == max(date)) %>%
  mutate(ranking = dense_rank(desc(Deaths))) %>%
   filter(ranking <= 10) %>% 
  pie_chart(group_var = country, title = "Top 10 Deaths by Country")

# pie of Recovered by country
pie_recovered <- recovered %>% filter(date == max(date)) %>%
  mutate(ranking = dense_rank(desc(Recovered))) %>%
   filter(ranking <= 10) %>% 
  pie_chart(group_var = country, title = "Top 10 Recovered by Country")

plot_grid(pie_confirmed, pie_deaths, pie_recovered, nrow = 3, labels = c('Confirmed', 'Deaths', 'Recovered'), 
          label_size = 12, scale = 1.5)


```
```{r eval=FALSE, include=FALSE}
## counts for Australia

confirmed %>%# select(-country) %>%
  filter(country=="Australia") %>%
  group_by(date) %>%
  summarise(Confirmed = sum(Confirmed)) %>%
  mutate(date = as.Date(date, format = "%m/%d/%y")) %>% # converts to a date class
  ggplot(aes(x = date, y = Confirmed)) + geom_line() +
      geom_col(fill = "orange") +
      labs(title = "Total Daily Confirmed Cases: Australia",
           subtitle = paste0(glue(dateforplot)),
           x = "Date", y = "Confirmed Cases") +
  theme_light()
  
deaths %>%# select(-country) %>%
  filter(country=="Australia") %>%
  group_by(date) %>%
  summarise(Deaths = sum(Deaths)) %>%
  mutate(date = as.Date(date, format = "%m/%d/%y")) %>% # converts to a date class
  ggplot(aes(x = date, y = Deaths)) + geom_line() +
      geom_col(fill = "orange") +
      labs(title = "Total Daily Deaths: Australia",
           subtitle = paste0(glue(dateforplot)),
           x = "Date", y = "Confirmed Deaths") +
  theme_light()

recovered %>%# select(-country) %>%
  filter(country=="Australia") %>%
  group_by(date) %>%
  summarise(Recovered = sum(Recovered)) %>%
  mutate(date = as.Date(date, format = "%m/%d/%y")) %>% # converts to a date class
  ggplot(aes(x = date, y = Recovered)) + geom_line() +
      geom_col(fill = "orange") +
      labs(title = "Total Daily Recoveries: Australia",
           subtitle = paste0(glue(dateforplot)),
           x = "Date", y = "Confirmed Recoveries") +
  theme_light()
      
```

## Appendix

```{r plotstesting1, eval=FALSE, include=FALSE}

all %>% 
  filter(group == "Confirmed") %>% # choose: "Confirmed" "Deaths"    "Recovered"
  filter(country==c("Australia", "Iran", "Italy", "US")) %>% # choose country
  group_by(country,date, group) %>% summarize(n = sum(n)) %>%
  ggplot(aes(x = date, y = n)) +
  geom_line(aes(color = country, linetype = country)) + 
  scale_color_manual(values = c("darkred", "steelblue", "red", "blue"))
# Try away from desktop due to wirte positions..
#
#p_load(gganimate)
#p <-all %>% 
  #filter(group == "Confirmed") %>% # choose: "Confirmed" "Deaths"    "Recovered"
  #filter(country==c("Australia", "Iran", "Italy", "US")) %>% # choose country
  #group_by(country,date, group) %>% summarize(n = sum(n)) %>%
  #ggplot(aes(x = date, y = n)) +
  #geom_point(aes(color = country)) + 
  #scale_color_manual(values = c("darkred", "steelblue", "red", "blue"))

#anim <- p + 
  #transition_states(country,
                    #transition_length = 2,
                    #state_length = 1)
#anim
#
```